Project Light

### 2015-09-09 paper/paper.tex
small fixes, add to conclusion

The Internet of Things (IoT) is becoming more and more reality every day.
It consists of small devices (called \textit{motes}), which are connected to the public Internet.
They work together, allowing monitoring and controlling of several kinds of real-world systems. % SOURCE!
E.g., motes collecting information about parking lot occupancy allowing car drivers to find free spaces using their smart phones\footnote{see http://www.streetline.com/}.

@@

In order to make the IoT reality, a specialized network stack is needed.
Motes have special constrains regarding energy and memory resources, e.g. batteries should work for multiple years and a mote's program memory will be very small (e.g. 48kb program flash for telosB mote, \cite{telosb}).
Additionally, typical IoT setups will make use of hundreds of nodes, so there many more addresses needed than the traditional Internet uses.

The \textit{IPv6 over Low-Power Wireless Personal Area Networks (6LoWPAN)} protocol stack as described in \cite{palattella2013standardized} is tailoring this needs.
The criteria for IoT network stacks are according to them: low power usage, high reliability, Internet-suitability.
Therefore, tailored implementations of stack layers are presented, like \textit{PHY of IEEE 802.15.4}, which is a power saving physical/link layer.
In order to be Internet-enabled regarding the need for huge amounts of addresses, IPv6 is utilized.
However, in order to save energy, the headers of IPv6 packets are heavily compressed in 6LoWPAN.

There are some proposals for specialized security measures in this stack, e.g. \textit{Adaptive Key Establish- ment Scheme (AKES)}, \cite{krentz15akes}.
AKES makes use of session keys and a neighborhood-exploration-technique to gain sufficient security level while being memory-space-saving.
Security is even obtained after reboots.
It still relies on predistribution keys on each mote.

The problem of wireless and user-friendly key predistribution / key provisioning remains.
There are multiple approaches to solve this:

\begin{itemize}
	\item \cite{chen2011over} proposes key provisioning using eliptic curve cryptography (ECC). However, while meeting high security standards, software implementations of ECC require a lot of memory space and energy while hardware implementations increase the per-unit cost.
	\item In order to secure radio communication, \cite{kuo2007message} utilizes a Faraday cage in which a mote can be programmed without interleaving of messages from an attacker outside of the cage. Additionally, a third device outside of the cave jams radio waves to prevent malicious communication.
	\item \textit{6doku} is a protocol for pairing  a mote with a touch-enabled devices like an smart phone, \cite{krentz20156doku}. The smart phone needs extra hardware attached for communicating with the mote via radio waves. It relies on interaction with the end-user, e.g. he needs to check if blinking on the screen happens in sync with him pressing a button on the mote.
\end{itemize}

@@

Project Light consists of an Android app that creates and sends the keying material and a Contiki process that runs on the TelosB sky mote and receives and validates it, \cite{dunkels04contiki}, \cite{telosb}.
Both follow the aforementioned communication protocol very closely, but increase reliability and fail-safeness through the following design decisions.

@@

The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smart phone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smart phone operating systems, too, e.g. iOS or Windows Phone 8.1.
In this version, almost all parameters for the transmission are adjustable by the user: length of one bit, length of initialization phase, initialization-pattern, start-pattern, if hamming-encoding should be used.
Therefore it is convenient to experiment with various settings and transmission parameters and to test of the Contiki process.
There are two disadvantages of this implementation, which motivated us to build a second edition:
First, experiments using the flash light for light emission showed, that there are no predictable timing behavior, e.g. the flash light was simply to slow to allow short transmission times.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Secondly, in order to get reliable transmission quality, we need to become as close to real-time as possible.
This technology stack indicates bad real-time behavior, i.g. the Cordova runtime and the JavaScript interpreter could be omitted.

% bis hier hin weg

\begin{figure}
	\centering
	\includegraphics[scale=.15]{images/screen_native.png}
	\caption{Screenshot of the native Android application used to encode the key and emit it as light flickering. With a toggle switch you can choose if you want the flashlight or the screen to flicker. The length of one bit can be adjusted using a slider. Any test entered in the text filed can be transmitted by tapping on ``start''. While transmitting, the user gets feedback about the current phase right above the buttons.}
	\label{fig:screenshot}
\end{figure}

% Und den dann natuerlich noch anpassen

Hence, we built a native Android application written in Java.
Still on Android, it is impossible to guarantee real-time behavior, but we may get smaller delays without JavaScript involved.
The Java implementation also benefits from the ability of the native Java timer to re-adjust itself, which means that timing errors do not add up.
In that way, the phone and the mote do not get out of sync, no matter how long the transmission takes.
% , e.g. if one event is fired too late such as wrong bit gets transmitted, the delay does not add up for the next bit.
This comes hand-in-hand with the Hamming(8,4)-encoding we used, because this could fix potential one-bit-errors in each byte.
Another advantage of this version is the flash light.
Depending on the smart phone model, the flash light was as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes very handy.
To improve usability and discoverability, we dropped controls for some parameters, which led to a much leaner design as shown in figure~\ref{fig:screenshot}.

However, as discussed in section~\ref{sec:results_and_discussion}, both versions worked well with similar values for the length of one bit.

@@

In this paper, we presented a secure method for IoT-device initialization by transmitting initial keying material via light.
It is mainly targeted at private users and can easily be executed having no technical background at all.

@@

To simplify the initialization of several motes at once, two things need be done.
First, the app should stay in calibration and synchronization phase, until the user presses a button.
Second, the mote should indicate via an LED that it reached the initialization phase~(\ref{sub:initialization}).
With those two concepts, the key transmission to several motes could work as follows:
The user lays out all motes in a grid next to each other.
Then the phone is placed on top of them and the app is started.
All motes are placed, so that they receive the flickering of the phones flash light.
The motes are started one after another.
Once the motes are running, the user waits for all of them to indicate via an LED, that they reached the initialization phase.
They are now all waiting for the initialization pattern.
The user presses a button in the app and the phone sends the initialization pattern followed by the data packet.
When the transmission is over, the user can see via a red LED which motes failed to receive the data correctly.
The process needs to be restarted with those motes.

% den Satz verstehe ich nicht
User-friendliness should be prioritized as success and adoption of this presented additional security measure will greatly benefit from it.

>>>>

The Internet of Things (IoT) is becoming more and more of a reality every day.
It consists of small devices (called \textit{motes}), which are connected to the public Internet.
They work together, allowing monitoring and controlling of several kinds of real-world systems. % SOURCE!
E.g., motes collecting information about parking lot occupancy allowing car drivers to find free spaces using their smart phones\footnote{see http://www.streetline.com/}.

@@

- IoT requires own network stack because of special needs: energy and resources, e.g. batteries should work for multiple years and mote memory will be very small (e.g. 48kb program flash for telosB mote CITE SOURCE), many addresses needed
- 6LowPan protocol stack as described in "Standardized protocol stack.." is tailoring this
- its using own implementations of layers, e.g. PHY for physical layer, and CoaP as applicationlayer as well as conventional ones (namely IPv6 as Internet layer) --> enables IoT

- there are some proposals for specialized security layers in this stack, e.g. akes \cite{krentz15akes}
- akes uses session keys and a neighborhood-exploration-technique to gain sufficient security level, even surviving reboots
- it still relies on predistributed keys on each mote
- how to predistribute in safe way?

\cite{chen2011over} proposes key provisioning without wiring using eliptic curve cryptography --> expensive for memory and energy.
- (own chips might solve this, but increase per unit cost)

\cite{kuo2007message} utilizes Faraday cage and another (third) radio jamming device for initialization.
-> requires special hardware

6doku \cite{krentz20156doku} proposes pairing with smartphone and extra-hardware for pairing
- extra hardware and motes need button

@@

Project Light consists of an Android app that creates and sends the keying material and a Contiki\cite{dunkels04contiki} process that runs on the TelosB sky mote and receives and validates it.
Both follow the aforementioned communication protocol very closely, but increase reliability and fail-safeness through the following design decisions.

@@

The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smart phone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smart phone operating systems, too, e.g. iOS or Windows Phone.
%In this version, almost all parameters for the transmission are adjustable by the user: length of one bit, length of initialization phase, initialization-pattern, start-pattern, if hamming-encoding should be used.
%Therefore it is convenient to experiment with various settings and transmission parameters and to test of the Contiki process.
There are two disadvantages of this implementation, which motivated us to build a second edition:
The flash light cannot be controlled reliably using the mentioned web technologies.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Furthermore the timing accuracy was considerably limited by the overhead caused by the Cordova runtime and the JavaScript interpreter which resulted in transmission errors.

% \begin{figure}
% 	\centering
% 	\includegraphics[scale=.35]{images/screenshots.png}
% 	\caption{The two Android App versions side-by-side (web on the left, native on the right).}
% 	\label{fig:screenshots}
% \end{figure}

To improve transmission reliability we therefore built a native Android application written in Java.
Still on Android, it is impossible to guarantee real-time behavior, but we may get smaller delays without JavaScript involved.
The Java implementation also benefits from the ability of the native Java timer to re-adjust itself, which means that timing inaccuracies do not add up.
In that way, the phone and the mote do not get out of sync, no matter how long the transmission takes.
% , e.g. if one event is fired too late such as wrong bit gets transmitted, the delay does not add up for the next bit.
%This comes hand-in-hand with the Hamming(8,4)-encoding we used, because this could fix potential one-bit-errors in each byte.
Another advantage of this version is the flash light.
Depending on the smart phone model, the flash light is as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes in very handy.
%To improve usability and discoverability, we dropped controls for some parameters, which led to a much leaner design as shown in figure~\ref{fig:screenshots}.

%However, as discussed in section~\ref{sec:results_and_discussion}, both versions worked well with similar values for the length of one bit.
% Ich glaub das diskutieren wir nicht, das passt iwie nicht zur Aussage des Papers

@@

In this paper, we presented a secure method for IoT-device initialization by transmitting initial keying material via light.
It is mainly targeted at private users and can easily be executed having no technical background at all.
As the visibility of light can be judged and restricted pretty well by end users, safety from interception is given.

@@

To simplify the initialization of several motes at once, two things need to be done.
First, the app should stay in calibration and synchronization phase, until the user presses a button.
Second, the mote should indicate via an LED that it reached the initialization phase~(\ref{sub:initialization}).
With those two concepts, the key transmission to several motes could work as follows:
The user lays out all motes in a grid next to each other.
Then the phone is placed on top of them and the app is started.
All motes are placed, so that they receive the flickering of the phones flash light.
The motes are started one after another.
Once the motes are running, the user waits for all of them to indicate via an LED, that they reached the initialization phase.
They are now all waiting for the initialization pattern.
The user presses a button in the app and the phone sends the initialization pattern followed by the data packet.
When the transmission is over, the user can see via a red LED which motes failed to receive the data correctly.
The process needs to be restarted with those motes.
